---
title: "Visualization"
subtitle: "Part I"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)

```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```


## Scatterplots!!

Create my first scatterplot ever. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
        geom_point()
```

aes matches which variables to which part of graph 
geom_point is for scatterplots

New approach, same plot

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax))+
        geom_point()
```
This was is a little more consistent with how we deal with data in tidy verse

save and edit a plot object.

```{r}
weather_plot =
        weather_df %>% 
        ggplot(aes(x=tmin, y=tmax))

weather_plot + geom_point()
```

## Advanced scatterplot...

Start with the same one and make it fancy

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax, color = name))+
        geom_point() + 
        geom_smooth(se = FALSE)
```

geom_smooth gives a smooth trajectory through our data

What about the aes placement?

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax))+
        geom_point(aes(color = name)) +
        geom_smooth()
```
Looks the same, until you add geom_smooth there are no other lines because color only applies to geom_point and not the gg_plot

Let's facet some things!!

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax, color = name))+
        geom_point(alpha = .2) + 
        geom_smooth(se = FALSE) +
        facet_grid(. ~ name )
```

Facet. first item defines rows and second defines columns, separated by tilde. 

alpha in geom_point is for transparency, can be defined by variable

Let's combine some elements and try a new plot

```{r}
weather_df %>% 
        ggplot(aes(x = date, y = tmax, color = name))+
        geom_point(aes(size = prcp),alpha = 0.5)+
        geom_smooth(se = FALSE)+
        facet_grid(. ~ name)
```



## Some small notes

How many geoms have to exist?

You can have whatever geoms you want 

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax, color = name))+
        geom_smooth(se = FALSE)
```

You can use a neat geom 
geom_hex for example

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = tmax))+
        geom_density2d()+
        geom_point(alpha = 0.3)
```

## Univariate plots
One variable at a time

Histograms are really great 

```{r}
weather_df %>% 
        ggplot(aes(x = tmin))+
        geom_histogram()
```

Can we add color to histograms?

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, fill = name))+
        geom_histogram()+
        facet_grid(. ~ name)
```

Use dodge to separate bars, fill fills in bars
        
Let's try a new geom

```{r}
weather_df %>% 
        ggplot(aes(x = tmin, fill = name))+
        geom_density(alpha = 0.4, adjust = .5)
```

Density is a histogram that's been smoothed out around the edges

What about box plots

```{r}
weather_df %>% 
        ggplot(aes(x = name, y = tmin))+
        geom_boxplot()
```

Trendy plots!

```{r}
weather_df %>% 
        ggplot(aes(x = name, y = tmin, fill = name)) +
        geom_violin(alpha = .5) +
        stat_summary(fun = "median")

```

default for stat_summary is the mean


Ride plots -- the most popular plot of 2017
```{r}
weather_df %>% 
        ggplot(aes(x = tmin, y = name))+
        geom_density_ridges()
```
















